# Tax Loss Harvesting

## Synopsis

## Data load and preprocessing
Load in the data from Yahoo and sort by date.
```{r, cache=TRUE}
library(dplyr)
library(ggplot2)

SPY <- read.csv('http://ichart.finance.yahoo.com/table.csv?s=SPY&a=1&b=1&c=2005&g=d&ignore=.csv')
SPY$Date <- as.Date(as.character(SPY$Date))
SPY <- arrange(SPY, Date)

ggplot(SPY, aes(x=Date, y=Close)) + geom_line()
```

## Buy and hold
We find the PNL for buy and hold. First we set the tax rates.
```{r}
capitalGainsTax <- 0.27
marginalTax <- 0.5
```

```{r}
library(dplyr)
library(ggplot2)
start <- SPY[1,]$Adj.Close

SPY <- (SPY %>% mutate(cumPNL=Close - start, taxadjustedPNL=start + pmax(0,cumPNL)*(1-capitalGainsTax) + pmin(0,cumPNL)*(1-marginalTax) ))

(ggplot(data=SPY) 
  + aes(Date)
  + geom_line(aes(y=Close, colour='raw')) 
  + geom_line(aes(y=taxadjustedPNL, colour='adjusted'))
  + ylab('SPY')
  + ggtitle('Buy and Hold')
  + scale_color_manual(name='Legend',values=c(raw='blue',adjusted='orange'))
 )
```

## Upper Bound for Tax Harvesting
In the following we assume we have an infinite number of securities with identical behaviour but which are treated as different for tax purposes. We employ a psychic person to trade at close with no transaction costs.

This serves to construct an upper bound for the benefits of tax harvesting.

### Split PNL into loss and gain
We first create a data frame with daily PNL deltas.
```{r}
dates <- SPY$Date
vals <- SPY$Close
deltas <- data.frame(
  Date=dates[2:length(dates)],
  Close=vals[2:length(dates)],
  prevClose=vals[1:length(dates)-1]
)
deltas$Delta <- deltas$Close - deltas$prevClose
```

```{r}
deltas <- select(deltas,Date,Delta)
SPY <- left_join(SPY, deltas, by='Date')
SPY[is.na(SPY$Delta),'Delta'] <- 0
SPY <- mutate(
    SPY, 
    cumGain=cumsum(pmax(0,Delta)), 
    cumLoss=cumsum(pmin(0,Delta)),
    psychic_taxadjustedPNL=start+(1-capitalGainsTax)*cumGain+(1-marginalTax)*cumLoss
  )
```

Plotting this upper bound:
```{r}
( ggplot(data=SPY)
  + aes(x=Date)
  + geom_line(aes(y=psychic_taxadjustedPNL, colour='psychic'))
  + geom_line(aes(y=Close, colour='Close'))
  + geom_line(aes(y=taxadjustedPNL, colour='B&H'))
  + ylab('PNL')
  + ggtitle('PNL by Strategy')
  + scale_color_manual(name='Legend',values=c(psychic='green',Close='orange', 'B&H'='violet'))
)
```